Abstractions for Multi-Sorted Substitutions

### 2023-09-16 paper/paper.tex
Update.

  With \emph{intrinsic typing}, the definitions of syntax and typing
  are merged into a single definition of well-typed terms.
  Consequently, operations on terms have to map
  well-typed terms to well-typed terms proving that the
  operations preserve typing.
  For example, the definition of substitution
  also acts as a proof that substitution preserves typing, and the
  definition of a small-step semantics also acts as a proof for subject
  reduction.
  This often has the benefit of more concise code, as shared structure
  between definitions and preservation lemmas does not need to be
  repeated.

  With \emph{extrinsic typing}, the definitions of syntax and typing
  are kept separately: the syntax is defined first, and then typing is
  defined as a relation. Consequently, operations on terms and
  their type preservation lemmas are also separate entities.
  While extrinsic typing cannot eliminate the shared structure between
  definitions and preservation lemmas, it can automated by reflection
  or generic programming techniques, leaving only the preservation lemmas
  to be proven by the programmer. While typing can be arbitrary complex,
  the definition and metatheory of untyped substitution behaves the same
  across languages.

>>>>

  With \emph{intrinsic typing}, the definitions of syntax and typing
  are merged into a single definition of well-typed terms.
  Consequently, the definition of an operation on terms also acts
  as a type preservation lemma for that operation, as it has to
  map well-typed terms to well-typed terms.
  For example, the definition of substitution
  also acts as a proof that substitution preserves typing, and the
  definition of a small-step semantics also acts as a proof for subject
  reduction.
  This often has the benefit of more concise code, as shared structure
  between definitions and preservation lemmas does not need to be
  repeated.

  With \emph{extrinsic typing}, the definitions of syntax and typing
  are kept separately: the syntax is defined first, and then typing is
  defined as a relation. Consequently, operations on terms and
  their type preservation lemmas are also separate entities.
  However, hope is not lost for extrinsic typing to keep up with the
  conciseness of intrinsic typing: 
  While extrinsic typing cannot eliminate the shared structure between
  definitions and preservation lemmas, it can automated by reflection
  or generic programming techniques, leaving only the preservation lemmas
  to be proven by the programmer. While typing can be arbitrary complex,
  the definition and metatheory of untyped substitution behaves the same
  across languages.
